\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-21 paper/paper.tex
updates to paper.

    // KV store core
    event_replicated_ganged_op (packet, barrier)
    {
      if(leader KV core)
        atomic set bit me in barrier.mask         
        do
         for each KV store core in packet
          if core_public_data[core].view > packet.view
           barrier.failed = true 
           atomic set bit for core in barrier.mask
        while barrrer.mask != mask of participating KV cores
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         core_public_data[leader core].view > packet.view
         OR core_public_data[me].barrier == mask of participating KV cores
        if core_public_data[leader core].view > packet.view
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

@@

\section{Related Work}
Memory technology in datacenters is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs~\cite{farm} as well as newer memory types such
as 3D XPoint~\cite{pmfs, bpfs}. In anticipation, libraries that allow
programmers to build durable data structures on non-volatile heaps have become
available~\cite{mnemosyne, nvheaps, cdds} - an example of such a library that we
have used in this paper is Intel's non volatile memory library
(NVML)~\cite{nvml}. A missing piece however is a component to make such durable
data structures highly available using replication across a commodity network
such as ethernet. Existing solutions that offer good performance are dependent
on high performance RDMA networking hardware, with a dependence on external
services (such as zookeeper~\cite{zookeeper}) and application specific code for
fault recovery~\cite{farm, htm}.  This is a problem for programmers using NVML
where one cannot apriori guarantee the availability of such RDMA networking
hardware (for e.g., when running on externally hosted environments or public
clouds).

\section{Conclusion}
Cyclone allows developers to add availability to key value maps built on top of
libraries such as NVML and achieve good performance on commodity networking
hardware. A key question is whether Cyclone can be used for other applications
such as filesystems? \emph{We believe that the requirement for scaling across
  multiple replication quorums can be satisfied by applications where the
  scalable commutativity rule~\cite{scalable_commutativity} applies to common
  case primitives.} Although our intitial release of Cyclone will be for key
value style applications, our on-going research is based on the scalable
commutativity hypothesis.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

    // KV store core
    event_replicated_ganged_op (packet, barrier)
    {
      if(leader KV core)
        atomic set bit me in barrier.mask         
        do
         for each KV store core in packet
          if core_public_data[core].view > packet.view
           barrier.failed = true 
           atomic set bit for core in barrier.mask
        while barrrer.mask != mask of participating KV cores
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         core_public_data[leader core].view > packet.view
         OR barrier.mask == mask of participating KV cores
        if core_public_data[leader core].view > packet.view
         OR barrier.failed
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

@@

\section{Related Work}

Cyclone takes a software only approach to improving replication performance over
commodity network hardware. At the other end of the spectrum, Consensus in a
box~\cite{consensus_box} implements the zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive from the FPGA that cuts
out the path from CPU caches to the NIC are as low as 3us for a round trip from
leader to follower replica compared to the 7us we observe with DPDK. Notably
however they reported 7us when using TCP/IP from the FPGA rather than their own
connection oriented network protocol. On the other hand peak replication
throughput appears to be 6M replications/sec on 66 Gbps of aggregate network
bandwith, comparable to the mark obtained by Cyclone on an aggregate 40 Gbps of
network bandwidth. We believe the general applicability afforded by our software
only packes switching solution makes it attractive in comparison to an FPGA
based solution even given the larger latency.

Networked Ordered Paxos~\cite{nopaxos} reprograms top of rack switches to order
client requests and multicast them to replicas - attacking replication latency
by removing the need to direct requests through a leader replica. It is hard to
compare performance due to the 1Gbps links used to leaf nodes in NoPaxos but we
note that avoiding changes to switching infrastructure was a specific constraint
in the design of Cyclone due to the difficulty we encountered in convincing
public cloud customers to allow such changes to be made.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL likely rendering
it unsuitable as a replacement for the WAL in key value stores.

A number of researchers have proposed using high performance networking hardware
for improving the performance of various distributed system stacks such as
distributed transactions~\cite{farm} and replication~\cite{dare, mojim}. We note
that the ideas in Cyclone are not specific to ethernet. The same ideas of packet
switching and horizontal scaling can be equally applied to RDMA. However, we
chose commodity ethernet due to the short term deployment plans that are a
priority design target.

The scalable commutativity rule~\cite{scalable_commutativity} generalized the
idea of commutativity in key value store interfaces that we have used for
horizontal scaling in Cyclone. Conveniently it provides a tool for checking when
API calls can commute in a history. It is a simple exercise to express a key
value store in COMMUTER and check that operations to different keys
commute. This provides a theoretical basis for horizontal scaling, something we
have omitted for brevity from this paper. The scalable commutativity rule also
suggests that scalable key value store implementations would exploit
commutativity something we have observed in the case of in-memory layers for key
value stores such as Flodb and indeed in the way we distributed keys across
physical logs in Cyclone.

Finally, we note the in progress transformation of durable memory in datacenters
with the advents of 3DXPoint. Researchers have speculated on the impact of cheap
and high volume directly attached 3DXPoint~\cite{tiering}. Memory vendors have
also begun shipping software libraries such as NVML~\cite{nvml} for building durable
in-memory data structures using undo logging. A durable key value store using
3DXPoint would be an idea target for Cyclone because each NVM log entry can be
discarded as soon as it is applied to the key value store, since the results are
persistent. This would allow Cyclone to operate at pure packet switching speeds,
a goal we hope to achieve in the future with a single level log for NVML
applications.

\section{Conclusion}

\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
